---
title: "Albert_Hakobyan2_DV_HW4R"
author: "Albert Hakobyan"
date: "2025-03-27"
output: html_document
---

```{r, echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
library(ggplot2)
library(dplyr)
library(ggthemes)
```

```{r}
bliga_df <- read.csv('bundesliga.csv')

head(bliga_df, 10)
```

## PART 4: Rivalries & Big Match Patterns (*R*):
> #### 1. Head-to-Head Matrix for Selected Rivalries
- Select 5 key rivalries more info click here .
- Create a facet grid of win/draw/loss bar charts per rivalry.
- Annotate biggest win margins.

```{r}
rivalries <- list("Bayern Munich vs Dortmund" = list(team1 = "Bayern Munich", team2 = "Dortmund"),
                  "Schalke 04 vs Dortmund" = list(team1 = "Schalke 04", team2 = "Dortmund"),
                  "Bayern Munich vs M'Gladbach" = list(team1 = "Bayern Munich", team2 = "M'Gladbach"),
                  "FC Koln vs M'Gladbach" = list(team1 = "FC Koln", team2 = "M'Gladbach"),
                  "Hamburg vs Werder Bremen" = list(team1 = "Hamburg", team2 = "Werder Bremen"))
```

```{r}
all_rivalry_matches <- NULL

for (rivalry_name in names(rivalries)) {
  team1 <- rivalries[[rivalry_name]]$team1
  team2 <- rivalries[[rivalry_name]]$team2
  
  matches <- bliga_df %>%
    filter((HOMETEAM == team1 & AWAYTEAM == team2) | (HOMETEAM == team2 & AWAYTEAM == team1)) %>%
    mutate(Rivalry = rivalry_name,
       Team1IsHome = HOMETEAM == team1,
            Result = case_when(
                        Team1IsHome & FTHG > FTAG ~ "Win",
                       !Team1IsHome & FTAG > FTHG ~ "Win",
                                     FTHG == FTAG ~ "Draw",
                                             TRUE ~ "Loss"),
    GoalDifference = ifelse(Team1IsHome, FTHG - FTAG, FTAG - FTHG))
  all_rivalry_matches <- bind_rows(all_rivalry_matches, matches)}

# Summary of wins/draws/losses
rivalry_summary <- all_rivalry_matches %>%
  group_by(Rivalry, Result) %>%
  summarise(Count = n(), .groups = "drop")

# Defining all_combinations (Ensuring existence)
all_combinations <- expand.grid(Rivalry = unique(all_rivalry_matches$Rivalry),
                                Result = c("Win", "Draw", "Loss"))

rivalry_summary <- left_join(all_combinations, rivalry_summary, by = c("Rivalry", "Result"))
rivalry_summary$Count[is.na(rivalry_summary$Count)] <- 0

# The biggest win margins 
biggest_wins <- all_rivalry_matches %>%
  filter(Result == "Win") %>%
  group_by(Rivalry) %>%
  slice_max(order_by = abs(GoalDifference), n = 1) %>%
  ungroup()

max_counts <- rivalry_summary %>% # Max counts per rivalry for annotation positioning
  group_by(Rivalry) %>%
  summarise(MaxCount = max(Count, na.rm = TRUE), .groups = "drop")

biggest_wins <- biggest_wins %>%
  left_join(max_counts, by = "Rivalry")

head(all_rivalry_matches)
```



```{r}
head(biggest_wins)
```



### Facet grid of bar charts
```{r fig.width=15, fig.height=10}
ggplot(rivalry_summary, aes(x = Result, y = Count, fill = Result)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Rivalry, ncol = 2) +
  geom_text(data = biggest_wins, aes(x = "Win", y = MaxCount * 1.1, label = paste("Biggest win:", abs(GoalDifference), "goals")), vjust = 0.3) +
  labs(title = "Head-to-Head Statistics for Selected Bundesliga Rivalries", x = "Result", y = "Number of Matches") +
  theme_bw() +
  scale_fill_manual(values = c("Win" = "gold", "Draw" = "grey", "Loss" = "red"))
```

> #### 2. Upset Visualizer
- Define “upset” as a team >8 places below beating a top-5 team.
- Scatterplot of upsets: x-axis = rank difference, y-axis = goal difference.
- Encode team colors; highlight and label famous upsets


```{r}
# First, we need to create a table of season standings to determine team rankings
standings <- bliga_df %>%
  group_by(SEASON, HOMETEAM) %>%
  summarise(
    Points = sum(case_when(
      FTHG > FTAG ~ 3,
      FTHG == FTAG ~ 1,
      TRUE ~ 0)),
    GoalsScored = sum(FTHG),
    GoalsConceded = sum(FTAG),
    .groups = "drop") %>%
  rename(TEAM = HOMETEAM)

# Then we should add "away" games to standings
away_standings <- bliga_df %>%
  group_by(SEASON, AWAYTEAM) %>%
  summarise(Points = sum(case_when(FTAG > FTHG ~ 3,
                                   FTAG == FTHG~ 1,
                                          TRUE ~ 0)),
                    GoalsScored = sum(FTAG),
                    GoalsConceded = sum(FTHG),
                    .groups = "drop") %>%
  rename(TEAM = AWAYTEAM)

head(away_standings)
```




```{r}
# Now we need the combining of home and away standings
total_standings <- bind_rows(standings, away_standings) %>%
  group_by(SEASON, TEAM) %>%
  summarise(Points = sum(Points),
    GoalsScored = sum(GoalsScored),
    GoalsConceded = sum(GoalsConceded),
    GoalDifference = GoalsScored - GoalsConceded,
    .groups = "drop")

head(total_standings)
```



```{r}
# We need to establish a "Rank" system of teams by points (primary) and goal difference (secondary) for each season
season_rankings <- total_standings %>%
                  group_by(SEASON) %>%
arrange(desc(Points), desc(GoalDifference), .by_group = TRUE) %>%
        mutate(Rank = row_number()) %>%
                  ungroup()

top_teams <- season_rankings %>%
  filter(Rank <= 5) %>% # Identifying top 5 teams per season (based on points)
  select(SEASON, TEAM)

head(top_teams, 10)
```

```{r}
upsets <- bliga_df %>%
  # Joining with rankings for home teams
  left_join(season_rankings, by = c("SEASON", "HOMETEAM" = "TEAM")) %>%
  rename(HomeRank = Rank) %>%
  # Joining with rankings for away teams
  left_join(season_rankings, by = c("SEASON", "AWAYTEAM" = "TEAM")) %>%
  rename(AwayRank = Rank) %>%
  # Filter for upsets
  filter(
    # Either home team is > 8 places below and beats a top5 away team
    ((HomeRank - AwayRank > 8) & (FTHG > FTAG) & (AwayRank <= 5)) |
    # Or away team is > 8 places below and beats a top5 home team
    ((AwayRank - HomeRank > 8) & (FTAG > FTHG) & (HomeRank <= 5))) %>%
  # Calculating thee rank difference and goal difference
  mutate(
    RankDifference = case_when(
      (HomeRank - AwayRank > 8) & (FTHG > FTAG) ~ HomeRank - AwayRank,
      (AwayRank - HomeRank > 8) & (FTAG > FTHG) ~ AwayRank - HomeRank,
      TRUE ~ NA_real_),
    GoalDifference = case_when(
      (HomeRank - AwayRank > 8) & (FTHG > FTAG) ~ FTHG - FTAG,
      (AwayRank - HomeRank > 8) & (FTAG > FTHG) ~ FTAG - FTHG,
      TRUE ~ NA_real_),
    UpsetTeam = case_when(
      (HomeRank - AwayRank > 8) & (FTHG > FTAG) ~ HOMETEAM,
      (AwayRank - HomeRank > 8) & (FTAG > FTHG) ~ AWAYTEAM,
      TRUE ~ NA_character_),
    FavoriteTeam = case_when(
      (HomeRank - AwayRank > 8) & (FTHG > FTAG) ~ AWAYTEAM,
      (AwayRank - HomeRank > 8) & (FTAG > FTHG) ~ HOMETEAM,
      TRUE ~ NA_character_),
    MatchLabel = paste0(UpsetTeam, " ", 
                        ifelse((HomeRank - AwayRank > 8) & (FTHG > FTAG), FTHG, FTAG), 
                        "-", 
                        ifelse((HomeRank - AwayRank > 8) & (FTHG > FTAG), FTAG, FTHG), 
                        " ", FavoriteTeam, " (", SEASON, ")"))

```

```{r}
head(upsets, 10)
```


```{r}
# Getting the famous upsets (top 3 by rank difference)
famous_upsets <- upsets %>%
  arrange(desc(RankDifference)) %>% head(3)

team_colors <- data.frame(
  TEAM = c(
    "Bayern Munich", "Dortmund", "Duisburg", "FC Koln", "Hamburg", 
    "Leipzig", "M'Gladbach", "Wattenscheid", "Werder Bremen", "Dresden", 
    "Ein Frankfurt", "Freiburg", "Kaiserslautern", "Karlsruhe", "Leverkusen", 
    "Nurnberg", "Schalke 04", "Stuttgart", "Uerdingen", "Bochum", 
    "Munich 1860", "M'gladbach", "Hansa Rostock", "St Pauli", "Dusseldorf", 
    "Bielefeld", "Hertha", "Wolfsburg", "Ulm", "Unterhaching", 
    "Cottbus", "Hannover", "Mainz", "Aachen", "Hoffenheim", 
    "Augsburg", "Greuther Furth", "Fortuna Dusseldorf", "Braunschweig", "Paderborn", 
    "Darmstadt", "Ingolstadt", "RB Leipzig"),
  Color = c(
    "#DC052D", "#FDE100", "#0046AD", "#ED1C24", "#0C2240", 
    "#B1003C", "#18A33C", "#F7D917", "#1D9053", "#F0E453", 
    "#E1000F", "#D31230", "#D3171E", "#0C4C92", "#E32221", 
    "#9B1C1F", "#004D9D", "#DA291C", "#E30613", "#144DA3", 
    "#006AB3", "#18A33C", "#00A5DC", "#A52A2A", "#EE1D23", 
    "#004F9F", "#004B9C", "#4C9E2F", "#001489", "#E30613", 
    "#E30613", "#C11D28", "#ED1C24", "#FFED00", "#1E87C5", 
    "#D80A14", "#0AAC4A", "#E4002B", "#FBBA00", "#005CA9", 
    "#0045A1", "#CC0033", "#B1003C"), stringsAsFactors = FALSE)

# Adding colors to upsets (default color for teams not in the list)
upsets <- upsets %>%
  left_join(team_colors, by = c("UpsetTeam" = "TEAM")) %>%
  mutate(TeamColor = ifelse(is.na(Color), "turquoise", Color))
```

```{r}
head(upsets, 10)
```


```{r fig.width=15, fig.height=10}
ggplot(upsets, aes(x = RankDifference, y = GoalDifference)) +
  geom_hline(yintercept = c(1, 2, 3, 4), linetype = "dashed") +
  geom_vline(xintercept = seq(9, 16, by = 1), linetype = "dashed") +
  
  geom_point(aes(color = UpsetTeam), size = 5, alpha = 0.8) +
  
  geom_point(data = famous_upsets, aes(x = RankDifference, y = GoalDifference), 
             size = 10, shape = 1, color = "turquoise", stroke = 1.5) +
  geom_text(data = famous_upsets, 
            aes(x = RankDifference, y = GoalDifference, label = MatchLabel),
            vjust = -1.5, hjust = 0.5, size = 3.5, fontface = "bold",
            check_overlap = TRUE) +
  scale_x_continuous(breaks = seq(9, 16, by = 1), 
                     minor_breaks = NULL,
                     limits = c(8.5, max(upsets$RankDifference) + 0.5)) +
  scale_y_continuous(breaks = seq(1, 4, by = 1),
                     minor_breaks = NULL,
                     limits = c(min(upsets$GoalDifference) - 0.2, max(upsets$GoalDifference) + 0.8)) +
  labs(title = "Bundesliga Upsets: Lower-Ranked Teams Beating Top 5 Teams",
       subtitle = "Upset definition: When a team ranked at least 9 places below beats a top-5 team", 
       x = "Rank Difference (9 = minimum qualifying difference, higher values = bigger upsets)", 
       y = "Goal Difference in Match",
       caption = "Turquoise colored circles highlight the most significant upsets (largest rank differences)") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold", size = 14),
        plot.subtitle = element_text(size = 12),
        legend.position = "right",
        legend.title = element_text(face = "bold"),
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.8, "lines"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "gray"),
        axis.title = element_text(face = "bold"),
        axis.text = element_text(size = 10),
        plot.caption = element_text(hjust = 0.5, size = 10, margin = margin(t = 10))) +
  scale_color_manual(values = setNames(team_colors$Color, team_colors$TEAM)) +
  guides(color = guide_legend(override.aes = list(size = 4), ncol = 1))
```

## PART 5: Overall performance
> Define unique color for each team per season. For each season create horizontal bar plot using total
number of points. Highlighting the winner with the unique color that you assigned to it. Save all
graphs in pdf. (*R*)

```{r}
seasons <- unique(bliga_df$SEASON)

pdf("bundesliga_season_standings.pdf", width = 20, height = 8)

for (season in seasons) {
  season_data <- bliga_df %>% filter(SEASON == season)
  
  home_points <- season_data %>%
    group_by(HOMETEAM) %>%
    summarise(Points = sum(case_when(FTHG > FTAG ~ 3, 
                                     FTHG == FTAG~ 1, 
                                            TRUE ~ 0))) %>%
    rename(Team = HOMETEAM)
  
  away_points <- season_data %>%
    group_by(AWAYTEAM) %>%
    summarise(Points = sum(case_when(FTAG > FTHG ~ 3, 
                                     FTAG == FTHG~ 1, 
                                            TRUE ~ 0))) %>%
    rename(Team = AWAYTEAM)
    # Combining points and sorting
  standings <- bind_rows(home_points, away_points) %>%
    group_by(Team) %>%
    summarise(TotalPoints = sum(Points)) %>%
    arrange(desc(TotalPoints))
  
  winner <- standings$Team[1]
  
  season_colors <- standings %>%
    left_join(team_colors, by = c("Team" = "TEAM")) %>%
    mutate(Color = ifelse(is.na(Color), "turquoise", Color)) # Default turquoise for missing teams
  
  color_values <- setNames(season_colors$Color, season_colors$Team) #colors for scale_fill_manual
  
  eachPlot <- ggplot(standings, aes(x = reorder(Team, TotalPoints), y = TotalPoints, fill = Team)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = TotalPoints), hjust = -0.2, size = 3) +
    geom_bar(data = filter(standings, Team == winner), 
             aes(x = reorder(Team, TotalPoints), y = TotalPoints),
             stat = "identity", fill = "gold", color = "black") +     # Highlighting the winner
    coord_flip() +
    scale_fill_manual(values = color_values) +
    labs(title = paste("Bundesliga Season", season), subtitle = paste("Winner:", winner), x = "", y = "Total Points") +
    theme_bw() +
    theme(legend.position = "none",
              plot.title = element_text(face = "bold"),
        panel.grid.minor = element_blank())
  print(eachPlot)}

dev.off()
```


## PART 6: Monte Carlo simulation. *R*
> *Use Monte Carlo simulation to predict how many goals will Bayern Munchen score for next 10 seasons. Re-
peat the same for Bayer Leverkusen and Borusia Dortmund. Compare results using appropriate visualization
technique.*

```{r}
# Function to calculate goals per season for a team
calculate_goals_per_season <- function(data, team_name) {
  home_goals <- data %>%
    filter(HOMETEAM == team_name) %>%
    group_by(SEASON) %>%
    summarise(HomeGoals = sum(FTHG))
  
  away_goals <- data %>%
    filter(AWAYTEAM == team_name) %>%
    group_by(SEASON) %>%
    summarise(AwayGoals = sum(FTAG))
  
  # Merging and calculating total goals per season
  full_join(home_goals, away_goals, by = "SEASON") %>%
    mutate(HomeGoals = ifelse(is.na(HomeGoals), 0, HomeGoals),
           AwayGoals = ifelse(is.na(AwayGoals), 0, AwayGoals),
          TotalGoals = HomeGoals + AwayGoals) %>%
    arrange(SEASON)}

# Extracting Historical goal data for each team
bayern_goals <- calculate_goals_per_season(bliga_df, "Bayern Munich")
dortmund_goals <- calculate_goals_per_season(bliga_df, "Dortmund")
leverkusen_goals <- calculate_goals_per_season(bliga_df, "Leverkusen")

calculate_team_stats <- function(goals_data) {
  mean_goals <- mean(goals_data$TotalGoals)
    sd_goals <- sd(goals_data$TotalGoals)
   min_goals <- min(goals_data$TotalGoals)
   max_goals <- max(goals_data$TotalGoals)
  
  return(list(
   mean = mean_goals,
     sd = sd_goals,
    min = min_goals,
    max = max_goals))}

bayern_stats <- calculate_team_stats(bayern_goals)
dortmund_stats <- calculate_team_stats(dortmund_goals)
leverkusen_stats <- calculate_team_stats(leverkusen_goals)
```

```{r}
print("Bayern Munich Statistics:")
print(bayern_stats)
print("Dortmund Statistics:")
print(dortmund_stats)
print("Leverkusen Statistics:")
print(leverkusen_stats)
```

### Monte Carlo Simulation Function
```{r}
monte_carlo_simulation <- function(team_stats, num_seasons = 10, num_simulations = 1000) {
  # Creating a matrix to store all simulations. Each row is a simulation, each column is a season
  simulation_results <- matrix(0, nrow = num_simulations, ncol = num_seasons)
  
  for (i in 1:num_simulations) {
    # According to step 1: I should generate random goals for each season based on historical pattern
    # Using a normal distribution with mean and standard deviation from historical data
    # I'll use the log-normal distribution because I should not allow negative goals
    
    # Turning mean and sd into log-normal parameters
    variance <- team_stats$sd^2
    mu <- log(team_stats$mean^2 / sqrt(variance + team_stats$mean^2))
    sigma <- sqrt(log(1 + (variance / team_stats$mean^2)))
    
    # Generating random goals for each season
    for (j in 1:num_seasons) {
      # Generating a random number from log-normal distribution
      random_value <- rlnorm(1, meanlog = mu, sdlog = sigma)
      
      # Rounding to nearest integer for goal count (Almost forgot about this ...)
      simulation_results[i, j] <- round(random_value)}
  }
  return(simulation_results)}
```


```{r}
set.seed(42)

# Now we need to set the number of simulations and future seasons we should step into. 
# P.S.: if this was not Monte Carlo, but a different time series forecasting method (such as AR[q], MA[p] or ARIMA/SARIMA), 
# where order of AR or MA is important (both seasonal and trend), I'd use an ACF or PACF and AIC tests to
# reliably determine the number of future seasons we could reliably select in our forecasting.
# But since this is MC, lets go with 10 steps into the future.

num_simulations <- 1000
future_seasons <- 10

# Performing Monte Carlo simulations for each team of interest
bayern_simulations <- monte_carlo_simulation(bayern_stats, future_seasons, num_simulations)
dortmund_simulations <- monte_carlo_simulation(dortmund_stats, future_seasons, num_simulations)
leverkusen_simulations <- monte_carlo_simulation(leverkusen_stats, future_seasons, num_simulations)
```


```{r}
# Setting up a function which calculates the summary statistics from simulations
calculate_simulation_summary <- function(simulations) {
  # Calculating mean, median, and percentiles for each season
  summary_data <- data.frame(
    Season = 1:ncol(simulations),
    Mean = apply(simulations, 2, mean),
    Median = apply(simulations, 2, median),
    Lower_CI = apply(simulations, 2, function(x) quantile(x, 0.025)),
    Upper_CI = apply(simulations, 2, function(x) quantile(x, 0.975)))
  return(summary_data)}

bayern_summary <- calculate_simulation_summary(bayern_simulations)
dortmund_summary <- calculate_simulation_summary(dortmund_simulations)
leverkusen_summary <- calculate_simulation_summary(leverkusen_simulations)

# Converting summaries to a combined format for visualization
bayern_summary$Team <- "Bayern Munich"
dortmund_summary$Team <- "Dortmund"
leverkusen_summary$Team <- "Leverkusen"
combined_summary <- rbind(bayern_summary, dortmund_summary, leverkusen_summary)

head(combined_summary)
```

### Creating a visualization of the simulation results
```{r fig.width=20, fig.height=8}
ggplot(combined_summary, aes(x = Season, y = Mean, color = Team, fill = Team)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = Lower_CI, ymax = Upper_CI), alpha = 0.2, linetype = 0) +
  labs(title = "Monte Carlo Simulation: Predicted Goals per Season",
    subtitle = "Based on 1000 simulations with 95% confidence interval", x = "Future Season", y = "Predicted Goals", caption = "Simulation based on historical Bundesliga data") +
  theme_bw() +
  scale_color_manual(values = c("Bayern Munich" = "red", "Dortmund" = "yellow", "Leverkusen" = "black")) +
  scale_fill_manual(values = c("Bayern Munich" =  "red", "Dortmund" = "yellow", "Leverkusen" = "black"))
```

### Creating a density plot to show the distribution of goals for the final season


```{r}
final_season <- data.frame(
  Goals = c(bayern_simulations[, future_seasons], 
            dortmund_simulations[, future_seasons], 
            leverkusen_simulations[, future_seasons]),
  Team = rep(c("Bayern Munich", "Dortmund", "Leverkusen"), each = num_simulations))

head(final_season)
```

```{r fig.width=20, fig.height=8}
ggplot(final_season, aes(x = Goals, fill = Team)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of Predicted Goals for Final Simulated Season", x = "Total Goals", y = "Density") +
  theme_bw() +
  scale_fill_manual(values = c("Bayern Munich" = "red", "Dortmund" = "yellow", "Leverkusen" = "black"))
```



### A rolling forecast visualization (I am not satisfied with the results but this is Monte Carlo, what more could I expect from this :) )
- Also, I am completely aware that there is a better way of creating a rolling forecast, where it is a good practice to shift the forecast a couple of seasons before the most recent value and having the comparison of each model's forecast line plots from the future with the actual data behavior as well (to calculate the errors and to later build bar plots of those MSE/MAE of each model compared to each other, but I thought of these things too late into the assignment completion and the deadline is too close for all of this). At least my knowledge is exposed through my words.

```{r}
rolling_forecast <- data.frame()

# Last historical season (for connecting historical data to forecast)
last_season <- max(bayern_goals$SEASON)

# Historical data for plotting
historical_data <- rbind(data.frame(Season = bayern_goals$SEASON, Goals = bayern_goals$TotalGoals, Team = "Bayern Munich"),
  data.frame(Season = dortmund_goals$SEASON, Goals = dortmund_goals$TotalGoals, Team = "Dortmund"),
  data.frame(Season = leverkusen_goals$SEASON, Goals = leverkusen_goals$TotalGoals, Team = "Leverkusen"))

head(historical_data)
```

### Future data for plotting
```{r}
future_data <- rbind(data.frame(Season = (last_season + 1):(last_season + future_seasons), 
              Goals = bayern_summary$Mean,
           Lower_CI = bayern_summary$Lower_CI,
           Upper_CI = bayern_summary$Upper_CI,
               Team = "Bayern Munich"),
  data.frame(Season = (last_season + 1):(last_season + future_seasons), 
              Goals = dortmund_summary$Mean,
           Lower_CI = dortmund_summary$Lower_CI,
           Upper_CI = dortmund_summary$Upper_CI,
               Team = "Dortmund"),
  data.frame(Season = (last_season + 1):(last_season + future_seasons), 
              Goals = leverkusen_summary$Mean,
           Lower_CI = leverkusen_summary$Lower_CI,
           Upper_CI = leverkusen_summary$Upper_CI,
               Team = "Leverkusen"))
head(future_data)
```


### Plotting historical data with forecast
```{r fig.width=20, fig.height=8}
ggplot() +
  # Historical data as points and lines (scatterplotted lineplot)
  geom_point(data = historical_data, aes(x = Season, y = Goals, color = Team), size = 2) +
  geom_line(data = historical_data, aes(x = Season, y = Goals, color = Team), size = 0.7) +
  
  # Future data as lines with ribbon for confidence interval
  geom_line(data = future_data, aes(x = Season, y = Goals, color = Team), size = 1, linetype = "dashed") +
  geom_ribbon(data = future_data, aes(x = Season, ymin = Lower_CI, ymax = Upper_CI, fill = Team), alpha = 0.2) +
  
  # Vertical line to separate historical from forecast (imitate a rolling forecast visualization)
  geom_vline(xintercept = last_season + 0.5, linetype = "dashed") +
  labs(title = "Historical Goals and Monte Carlo Forecast (Rolling Horizon)", subtitle = "Historical data with 10-season forecast and 95% confidence intervals", x = "Season", y = "Total Goals", caption = "Dashed line separates historical data from forecast") +
  theme_solarized() +
  scale_color_manual(values = c("Bayern Munich" = "red",  "Dortmund" = "yellow", "Leverkusen" = "black")) +
  scale_fill_manual(values = c("Bayern Munich" =  "red",  "Dortmund" = "yellow", "Leverkusen" = "black"))
```


### Prediction summaries
```{r}
print("Bayern Munich Prediction Summary:")
print(bayern_summary)
print("Dortmund Prediction Summary:")
print(dortmund_summary)
print("Leverkusen Prediction Summary:")
print(leverkusen_summary)
```


### Creating bundesliga3.csv from bundesliga.csv to replicate bundesliga2.csv.
```{r}
bundesliga <- read.csv("bundesliga.csv")
transform_to_standings <- function(match_data) {
  home_stats <- match_data %>%
    group_by(SEASON, HOMETEAM) %>%
    summarise(Home_M = n(),  # Number of home matches
              Home_W = sum(FTHG > FTAG),       # Home wins
              Home_D = sum(FTHG == FTAG),      # Home draws
              Home_L = sum(FTHG < FTAG),       # Home losses
             Home_GF = sum(FTHG),             # Goals scored at home
             Home_GA = sum(FTAG),             # Goals conceded at home
         Home_Points = sum(FTHG > FTAG) * 3 + sum(FTHG == FTAG) * 1) %>%  # Points earned at home
         rename(TEAM = HOMETEAM)
  
  away_stats <- match_data %>%
    group_by(SEASON, AWAYTEAM) %>%
    summarise(Away_M = n(),                                 # Number of away matches
              Away_W = sum(FTAG > FTHG),
              Away_D = sum(FTAG == FTHG),
              Away_L = sum(FTAG < FTHG),                  
             Away_GF = sum(FTAG),                        
             Away_GA = sum(FTHG),                        
         Away_Points = sum(FTAG > FTHG) * 3 + sum(FTAG == FTHG) * 1) %>%  # Points earned away
         rename(TEAM = AWAYTEAM)
  
  # Step 3: Combine home and away statistics
  combined_stats <- full_join(home_stats, away_stats, by = c("SEASON", "TEAM")) %>%
    mutate(
      # Calculate total statistics
            M = Home_M + Away_M,              # Total matches
            W = Home_W + Away_W,              # Total wins
            D = Home_D + Away_D,              # Total draws
            L = Home_L + Away_L,              # Total losses
           GF = Home_GF + Away_GF,           # Total goals for
           GA = Home_GA + Away_GA,           # Total goals against
         DIFF = (Home_GF + Away_GF) - (Home_GA + Away_GA),  # Goal difference
       POINTS = Home_Points + Away_Points)  # Total points
  
  standings <- combined_stats %>%
    group_by(SEASON) %>%
    arrange(desc(POINTS), desc(DIFF), desc(GF), .by_group = TRUE) %>%
    mutate(POSITION = row_number()) %>%
    ungroup() %>%
    # Selecting only the columns needed for the final output
    select(TEAM, M, W, D, L, GF, GA, DIFF, POINTS, POSITION, SEASON)
  
  return(standings)
}
# Applying the transformation and saving the result
bundesliga_standings <- transform_to_standings(bundesliga)
# THE FILE IS called:  bundesliga3.csv
write.csv(bundesliga_standings, "bundesliga3.csv", row.names = FALSE)
head(bundesliga_standings)
```